Landau quantization of nearly degenerate bands

### 2018-09-20 paper.tex
Updates from ShareLaTeX

To the first order, $\delta \hat{H}$ and the Zeeman interaction induce a subleading phase offset ($\lambda_a$) which generically differs for the two orthogonal spinor states (labelled $a{=}1,2$):
\begin{equation}
l^2S(E)+\lambda_a(E,B)+\gamma=2\pi n,\as  n\in \mathbb{Z},\;a=1,2. \label{eq:rule}
\end{equation}
This implies that the spin-degeneracy of Landau levels is split   by approximately  $\var_c|\lambda_1{-}\lambda_2|/2\pi$. The phase offsets $\lambda_a$ can be viewed as Floquet quasienergies of a time-periodic, perturbative Hamiltonian ($\calh$) that governs the dynamics of the spinor wavefunction.  $\lambda_a$ is obtained  from  eigenvalues $\{e^{i\lambda_a}\}_{a=1}^2$ of a time-evolution propagator -- the time-ordered exponential 
\begin{equation}
\A{:}{=}\overline{\exp}\left[-\frac{i}{\hbar}\int_0^{T_c} \calh(\bk(t)) dt\right],
\label{eq:prop}
\end{equation} 
with $\bk(t){\in}\frako_0$ a periodic function determined by the semiclassical equation of motion.   $\calh$ has the two-by-two matrix form:
\begin{equation}
  \calh=\delta \var +B\bigg(M_z -g_s\mu_{B}\f{s_z}{\hbar} + e\epsilon_{\alpha\beta}\mathfrak{X}_{\beta}v_{\alpha}\bigg):=\delta \var+B\calm,\label{eq:H1}
\end{equation}
$\delta \var$ is the projection of $\delta H$ to the quasidegenerate band associated to $\var(\bk)$, e.g., $\delta \var{=}\alpha(k_x\sigma_y{-}k_y\sigma_x)$ in the Rashba model. In the basis of energy-split bands (which we label by ${+}$ and $-$ for bands corresponding to the outer and inner orbit respectively),\footnote{We remark that the quantization rule is invariant under basis transformations within the degenerate band subspace (See App. \ref{app:quantizationruleproof}); this may be exploited for efficient numerical computations.} $\delta \var$ is a diagonal matrix whose diagonal entries ($\delta \var_+,\delta \var_-$, corresponding to outer and inner orbit respectively; $\delta \var_+<\delta \var_-$) are related to the hybridization parameter:\footnote{The neglected term on the right-hand side of  \q{spinsplitwkb} is smaller (than the kept term) by a factor $\delta v_{\pm}/v$, which is the relative change in band speed $v=|\boldsymbol{v}|=\hbar^{-1}|\partial \var/\partial \bk|$ induced by the spin-orbit coupling. This factor is assumed to be small.  For example in the Rashba model, $\delta v_{\pm}/v=m\alpha/\hbar k_E{\sim}\delta S/S{\ll}1$.}   
\e{-\int_0^{T_c} (\delta \var_+{-}\delta \var_-) \frac{dt}{\hbar} \approx l^2\delta S. \la{spinsplitwkb}}
%In the  Rashba model,   $\delta \var_{\pm}(E){=}{\pm} (\alpha/\hbar)\sqrt{2mE}$ is independent of $\bk$ due to a continuous rotational symmetry of $H_R$. \\

@@

The remaining terms in \q{eq:H1} describe a generalized Zeeman interaction $B\calm$ which is the first-order-in-$B$ correction to the Peierls-Onsager Hamiltonian $\var(\bK)$.[kohn,blount,roth] Included in this interaction is the Zeeman coupling to the spin ($g_s\mu_Bs_z/\hbar$) and orbital ($M_z$) magnetic moment of the quasidegenerate band; the latter is expressible in terms of matrix elements of the velocity operator $\boldsymbol{\Pi}_{mn}=i\langle\psi_m|[\hat{H}, \hat{\boldsymbol{r}}]|\psi_n\rangle/\hbar$ between Bloch wave functions $\psi_{n\bk}$:
\e{
M_z=\f{ie\hbar}{2}\sum_l'\f{\Pi_{x,ml}\Pi_{y,ln}-\Pi_{y,ml}\Pi_{x,ln}}{\epsilon_m-\epsilon_l}. \la{eq:orbitalmoment}
}
Here, the summation in $l$ does not include the two quasidegenerate bands.  There is finally a geometric contribution to the Zeeman interaction, which involves the two-by-two non-Abelian Berry connection  as well as the  zeroth-order band velocity: 
\e{\mathfrak{X}_{\beta,mn}{=}i\braket{u_m}{\partial u_n/\partial k_{\beta}},  \;\;v_{\alpha}:= \red{\f{1}{\hbar}}\p{\var}{k_{\alpha}},\la{berryconn}}
with $\{u_n\}_{n=1}^2$ the periodic component of Bloch functions in the quasidegenerate band. 
%which provides a geometric characterization of band wavefunctions along $\frako_0$.\\
%All matrix elements in \qq{eq:prop}{eq:orbitalmoment} are evaluated with wavefunctions of $\hat{H}_0$.

\qq{eq:rule}{berryconn} allow us to determine $\lambda_{1,2}$ to the lowest order  in $1/l^2S$ and $\delta S/S$, and for arbitrary values of $l^2\delta S$ and $|B\calm|/\delta \var_{\pm}$.

@@

For simplicity, we assume that the Pauli matrices in \q{eq:Rashba-Hamiltonian} correspond to spin: $s_i{=}\hbar \sigma_i/2$. Let us evaluate the effective Hamiltonian $\calh(\bk)$  along the cyclotron trajectory $\frako_0$ with radius $k$. In the  energy basis of the Rashba Hamiltonian $H_R$ [cf.\ \q{eq:Rashba-Hamiltonian}], 
\e{\calh(\bk)=-\alpha k\tau_z{+}\f{g_s}{2}\mu_{\sma{B}}B\tau_x {-} \f{eB\hbar}{2m} (\tau_x-I) \epsilon_{\alpha\beta} k_{\alpha}\nabla_{\beta}\theta,\la{rashbaeffham}}
where $\tau_i$ are a new set of Pauli matrices with $\tau_z{=}{+} 1$ and ${-}1$ corresponding to the outer and inner orbits, respectively. $e^{i\theta_{\bk}}{=}(ik_x{-}k_y)/k$ is the angle of the in-plane spin-orbit field $({-}\alpha k_y,\alpha k_x,0)$, which winds once over $\frako_0$; the spin of our basis vectors likewise winds, in parallel (or anti-parallel) alignment with this field (see Fig. (\ref{fig:orbits})(b)). Terms proportional to the rate of spin rotation  ($\nabk \theta$) originate from the  non-abelian Berry connection. The first term of  $\calh$ corresponds to the spin-orbit-induced splitting of the energy-degenerate bands;
the remaining terms may be interpreted as B-induced deformations of the bands.   It is well-known that the Zeeman term tends to polarize spin in the direction ($\vec{z}$) of $\bB$; remarkably, the non-abelian Berry term   also has a polarizing effect, but with opposite sign. These two effects completely cancel for free electrons ($m{=}m_0$, with $m_0$ being free electron mass.), but not generically in band systems. The orbital moment term $B M_z$, which generally couples the two-band subspace (of interest) to all other bands, is absent in any simplified two-band model.
%Observe that  that $\calh$ is actually time-independent, which originates from a  that is unbroken by the field (aligned with the rotational axis). Consequently,

@@

Our quantization rule equally has utility where no closed-form analytic expression exists for the spectrum. Such is the case when we introduce a Dresselhaus spin-orbit coupling:
\begin{equation}
 H_{RD}(\bk)=\frac{{\hbar^2}k^2}{2m}+\alpha(k_{x}\sigma_{y}-k_{y}\sigma_{x})+\beta(k_{x}\sigma_{x}-k_{y}\sigma_{y}).\la{hamRD}
\end{equation}
In Fig. \ref{fig:RD}(b), a comparison is made between the Landau levels obtained from exact numerical diagonalization of $H_{RD}(\bK)$ (by standard techniques reviewed in App. \ref{sec:numerical}) and those obtained from the quantization rule. The difference in energies  does not exceed $2\%$ of $\var_c$ [cf.\ \fig{fig:RD}(c)] for various choices of the spin-orbit parameters $(\alpha,\beta)$.
Larger differences are positively correlated with 
$l^2\delta S^2/S$ (with $\delta S$ depending on both $\alpha$ and $\beta$), as shown in Fig. \ref{fig:RD}(d); see also \q{eq:Rashba-exact}, which analytically demonstrates how terms involving $l^2\delta S^2/S$ modifies the quantization rule in the case $\beta=0$.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  {spin-orbit field} ($d_x,d_y$) of the Hamiltonian $H_{RZ}{=}d_0(\bk)I{+}d_x(\bk)\sigma_x{+}d_y(\bk)\sigma_y$, with $d_j$ obtainable from \q{eq:RZ-Hamiltonian}. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-orbit field. (b) Black dashed lines illustrate a zeroth-order approximation of the Landau levels (assuming Klein tunneling); the leading-order correction is indicated by red lines. The Dirac point lies at $10$meV. Parameters for (b): $m$: 0.02$\hbar^2$/eV$\cdot$\AA$^2$, $\alpha$: 0.05eV$\cdot$\AA, $Z_\parallel$: 0.001eV. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4<\epsilon_0<1.0$(eV).\footnote{$l^2S\gg 1$ is assured by requiring $\epsilon_0>0.4$; the reason of choosing $\epsilon_0<1.0$ is Dirac point lying at high energy requires too much harmonics in numerical diagonalization. This requirement is equivalent to $\sqrt{0.8m}\alpha/\hbar<Z<\sqrt{2m}\alpha/\hbar$, corresponding to the boundaries of the plots in (d).} Parameters for (d): $m$: 0.01$\hbar^2$/eV$\cdot$\AA$^2$, $l=100$\AA. \textbf{[please choose formatting such that footnote is at the back]}\red{[Sorry, I haven't figured out how to do this. I will do this later.]} \label{fig:RZ}}
\end{figure}

@@

As alluded to in \s{sec:qtznrules}, the adiabatic limit of two independent orbits does not exist if the two spin-orbit-split orbits exactly intersect at a II-Dirac point. To investigate this failure of adiabaticity, let us study the Landau levels at energies close to the II-Dirac point ($\epsilon_0$). We focus first on the regime where, on average (over $\frako_0$), the energy splitting $|\delta \var_+{-}\delta \var_-|$ (due to both  spin-orbit and in-plane Zeeman interactions) is much greater than the interband, out-of-plane Zeeman interaction $B_{\perp}|\calm_{\pm}|$. This implies that the split orbits do not appreciably hybridize -- except in the vicinity of the II-Dirac point where orbits (nearly) touch. Near this II-Dirac point, the dynamics of a two-component wavepacket is described 
 by the effective Hamiltonian
\e{
\calh = \alpha k_{E} \,\omega^{\sma{\perp}}_c t\,\sigma_x+(Z_\parallel-\alpha k_{E})\sigma_y-Z_\perp\sigma_z,\la{effhamIIDirac}
}
with $k_E{=}\sqrt{2mE}/\hbar$ and  $t$ a time variable which parametrizes the zeroth-order orbit:
$\bk(t){=}({-}k_E \cos \omega_c^{\sma{\perp}} t, k_E \sin \omega_c^{\sma{\perp}} t ).$ 
$\calh$ above has been linearized with respect to  $t$, which vanishes in the vicinity of the Dirac point. The eigenbasis of $\sigma_x$ corresponds to two diabatic levels with characteristic slope $v_d{:}{=}\alpha k_E\omega^{\sma{\perp}}_c$; near $t{=}0$, the diabatic levels anticross  due to the hybridization terms proportional to $\sigma_{y,z}$. Generally, the probability of tunneling across the hybridization-induced barrier is  $\rho^2{=}\exp(-2\pi\barmu)$, with $\barmu{=}{E_g}^2/{2v_d \hbar}$ and $E_g$ the energy gap at the center of the anticrossing;[Landau and lifshitz, Wittig] in this context,
\e{\barmu=\frac{Z_\perp^2+(Z_\parallel-\alpha k_{E})^2}{2\alpha k_{E}\var^{\sma{\perp}}_c}, \la{mu1}}
with $2\alpha k_E$ the spin-orbit splitting at energy $E$ [cf.\ \fig{fig:orbits}(a)]. In the absence of the out-of-plane Zeeman coupling ($Z_\perp{=}0$), $\bar{\mu}$ would vanish at the energy of the II-Dirac point -- leading to unit-probability Klein tunneling, as was first proposed by [Obrien] in a different model realization of the II-Dirac point. However, with a proper account of $Z_\perp$, we find instead that $\bar{\mu}$ has a nonvanishing minimum: $\bar{\mu}_{\sma{\text{min}}}{:}{=} ( m/m_0)(Z_\perp/4\alpha k_{E})$, which is a product of the dimensionless effective mass and the ratio of the out-of-plane Zeeman splitting over the spin-orbit splitting. This minimum is practically negligible for  semiconductor heterostructures with $m{\ll}m_0$, but may be relevant in heavy-fermion systems.

To determine the Landau levels, one needs not just the probability for tunneling but also the phase of the probability amplitude. The full information is encoded in a scattering matrix [AALG]
\e{\mathbb{S}=\matrixtwo{\tau e^{i\barphi}}{-\rho}{\rho}{\tau e^{-i\barphi}},\;\rho=e^{-\pi\barmu},\;\tau=\sqrt{1-\rho^{2}}\la{scattmat}}
which is a connection formula that matches the two-component semiclassical (WKB) wavefunction across the Dirac point. $\mathbb{S}$ in \q{scattmat} has been expressed in the basis of Zeeman coupling corrected energy bands, i.e., eigenvector of Eq. \ref{effhamIIDirac}, so the diagonal elements of $\mathbb{S}$ are amplitudes for intraband (adiabatic) transmission; the phase of this amplitude is $\barphi{:}{=}\barmu{-}\barmu\ln\barmu{+}\text{arg}[\Gamma(i\barmu)]{+}\pi/4$ with $\Gamma$ the gamma function.

Away from the II-Dirac point, a semiclassical wavepacket undergoes adiabatic (band-conserving) dynamics. The time-ordered, coherent process of interband tunneling and adiabatic dynamics is described by the propagator:
\e{{\A}= \matrixtwo{\tau e^{i\bar{\varphi}}}{-\rho}{\rho}{\tau e^{-i\bar{\varphi}}} \diagmatrix{e^{i\tilde{\lambda}_+}}{e^{i\tilde{\lambda}_-}}. \la{propinplanezeeman}}
In the basis of Zeeman coupling corrected energy bands, the band-conserving phases $\tilde{\lambda}_{\pm}$ consist of two parts: dynamical phase $l^2 \delta S$ with $\delta S$ now accounting for both spin-split and Zeeman-split area corrections, and single band Berry phases $\phi^B_\pm$ evaluated on $\frako_0$. At energies far above the Dirac point ($|E{-}\epsilon_0|{\gg}|Z_{\perp}|$) where the Zeeman interaction is negligible, both orbits encircle the Dirac point and hence $\phi_{\pm}^B{\approx}{\pm}\pi$; far below the Dirac point, neither orbit encircles the Dirac point, leading to $\phi_{\pm}^B{\approx}0$. At intermediate energies, $\phi_{+}^B$ (resp.\ $\phi_{-}^B$) is a continuous and increasing (resp.\ decreasing) function of energy, as supported by the Bloch-sphere argument in Fig. \ref{fig:blochsphere}.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{blochsphere.png}
\caption{Energy dependence of the Berry phases for inner (upper panel) and outer (lower panel) orbits. For two band models, the Berry phase equals half the solid angle enclosed by the loop swept by spin-orbit field on the unit sphere. The solid angles are indicated by blue outlines and the energy is measured from $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}).\label{fig:blochsphere}}
\end{figure}

Our quantization rule [\qq{eq:rule}{eq:H1}], with $\A$ having the form of \q{propinplanezeeman}, may be equivalently expressed as
\e{
\text{cos}\left[\frac{\Omega_{-}+\Omega_{+}}{2}\right]=\tau\,\text{cos}\left[\frac{\Omega_{-}-\Omega_{+}}{2}+\bar{\varphi}\right], \lin
\ins{with}\;\Omega_{\pm}=l^2 (S\pm \delta S/2)+\phi_{\pm}^B +\gamma_{\pm}, \la{coscos}
}
$\Omega_{\pm}$ is the phase acquired by an electron in adiabatically traversing orbit $\pm$; the quantum Maslov correction $\gamma_{\pm}{=}\pi$ for circular orbits. Since $\phi^B_\pm$ are evaluated both on $\frako_0$, $\phi^B_-+\phi^B_+=0$ in two band models. A quantization rule analogous to \q{coscos} was previously derived by us for a II-Dirac point that intermediates an electron- and hole-like orbit [AALG].

>>>>

To the first order, $\delta \hat{H}$ and the Zeeman interaction induce a subleading phase offset ($\lambda_a$) which generically differs for the two orthogonal spinor states (labelled $a{=}1,2$):
\begin{equation}
l^2S(E)+\lambda_a(E,B)+\gamma=2\pi n,\as  n\in \mathbb{Z},\;a=1,2. \label{eq:rule}
\end{equation}
This implies that the spin-degeneracy of Landau levels is split   by approximately  $\var_c|\lambda_1{-}\lambda_2|/2\pi$. The phase offsets $\lambda_a$ can be viewed as Floquet quasienergies of a time-periodic, perturbative Hamiltonian ($\calh$) that governs the dynamics of the spinor wavefunction.  $\lambda_a$ is obtained  from  eigenvalues $\{e^{i\lambda_a}\}_{a=1}^2$ of a time-evolution propagator -- the time-ordered exponential 
\begin{equation}
\A{:}{=}\overline{\exp}\left[-\frac{i}{\hbar}\int_0^{T_c} \calh(\bk(t)) dt\right],
\label{eq:prop}
\end{equation} 
with $\bk(t){\in}\frako_0$ a periodic function determined by the semiclassical equation of motion.   $\calh$ has the two-by-two matrix form:
\begin{equation}
  \calh=\delta \var +B\bigg(M_z -g_s\mu_{B}\f{s_z}{\hbar} + e\epsilon_{\alpha\beta}\mathfrak{X}_{\beta}v_{\alpha}\bigg):=\delta \var+B\calm,\label{eq:H1}
\end{equation}
$\delta \var$ is the projection of $\delta H$ to the quasidegenerate band associated to $\var(\bk)$, e.g., $\delta \var{=}\alpha(k_x\sigma_y{-}k_y\sigma_x)$ in the Rashba model. In the basis of energy-split bands (which we label by ${+}$ and $-$ for bands corresponding to the outer and inner orbit respectively), $\delta \var$ is a diagonal matrix whose diagonal entries ($\delta \var_+,\delta \var_-$, corresponding to outer and inner orbit respectively; $\delta \var_+<\delta \var_-$) are related to the hybridization parameter:\footnote{The neglected term on the right-hand side of  \q{spinsplitwkb} is smaller (than the kept term) by a factor $\delta v_{\pm}/v$, which is the relative change in band speed $v=|\boldsymbol{v}|=\hbar^{-1}|\partial \var/\partial \bk|$ induced by the spin-orbit coupling. This factor is assumed to be small.  For example in the Rashba model, $\delta v_{\pm}/v=m\alpha/\hbar k_E{\sim}\delta S/S{\ll}1$.}   
\e{-\int_0^{T_c} (\delta \var_+{-}\delta \var_-) \frac{dt}{\hbar} \approx l^2\delta S. \la{spinsplitwkb}}
%In the  Rashba model,   $\delta \var_{\pm}(E){=}{\pm} (\alpha/\hbar)\sqrt{2mE}$ is independent of $\bk$ due to a continuous rotational symmetry of $H_R$. \\

@@

The remaining terms in \q{eq:H1} describe a generalized Zeeman interaction $B\calm$ which is the first-order-in-$B$ correction to the Peierls-Onsager Hamiltonian $\var(\bK)$.[kohn,blount,roth] Included in this interaction is the Zeeman coupling to the spin ($g_s\mu_Bs_z/\hbar$) and orbital ($M_z$) magnetic moment of the quasidegenerate band; the latter is expressible in terms of matrix elements of the velocity operator $\boldsymbol{\Pi}_{mn}=i\langle\psi_m|[\hat{H}, \hat{\boldsymbol{r}}]|\psi_n\rangle/\hbar$ between Bloch wave functions $\psi_{n\bk}$:
\e{
M_z=\f{ie\hbar}{2}\sum_l'\f{\Pi_{x,ml}\Pi_{y,ln}-\Pi_{y,ml}\Pi_{x,ln}}{\epsilon_m-\epsilon_l}. \la{eq:orbitalmoment}
}
Here, the summation in $l$ does not include the two quasidegenerate bands.  There is finally a geometric contribution to the Zeeman interaction, which involves the two-by-two non-Abelian Berry connection  as well as the  zeroth-order band velocity: 
\e{\mathfrak{X}_{\beta,mn}{=}i\braket{u_m}{\partial u_n/\partial k_{\beta}},  \;\;v_{\alpha}:= \f{1}{\hbar}\p{\var}{k_{\alpha}},\la{berryconn}}
with $\{u_n\}_{n=1}^2$ the periodic component of Bloch functions in the quasidegenerate band. 
%which provides a geometric characterization of band wavefunctions along $\frako_0$.\\
%All matrix elements in \qq{eq:prop}{eq:orbitalmoment} are evaluated with wavefunctions of $\hat{H}_0$.

\qq{eq:rule}{berryconn} allow us to determine $\lambda_{1,2}$ to the lowest order  in $1/l^2S$ and $\delta S/S$, and for arbitrary values of $l^2\delta S$ and $|B\calm|/\delta \var_{\pm}$. We remark that the quantization rule is invariant under basis transformations within the degenerate band subspace (See App. \ref{app:quantizationruleproof}); other basis, such as $\bk$-independent basis where Berry connections are trivial, may be more convenient in many discussions.

@@

For simplicity, we assume that the Pauli matrices in \q{eq:Rashba-Hamiltonian} correspond to spin: $s_i{=}\hbar \sigma_i/2$. Let us evaluate the effective Hamiltonian $\calh(\bk)$  along the cyclotron trajectory $\frako_0$ with radius $k$. In the  energy basis of the Rashba Hamiltonian $H_R$ [cf.\ \q{eq:Rashba-Hamiltonian}], 
\e{\calh(\bk)=-\alpha k\tau_z{+}\f{g_s}{2}\mu_{\sma{B}}B\tau_x {-} \f{eB\hbar}{2m} (\tau_x-I) \epsilon_{\alpha\beta} k_{\alpha}\nabla_{\beta}\theta,\la{rashbaeffham}}
where $\tau_i$ are a new set of Pauli matrices with $\tau_z{=}{+} 1$ and ${-}1$ corresponding to the outer and inner orbits, respectively. $e^{i\theta_{\bk}}{=}(ik_x{-}k_y)/k$ is the angle of the in-plane spin-splitting field $({-}\alpha k_y,\alpha k_x,0)$ of $H_R$, which winds once over $\frako_0$; the spin of our basis vectors likewise winds, in parallel (or anti-parallel) alignment with this field (see Fig. (\ref{fig:orbits})(b)). Terms proportional to the rate of spin rotation  ($\nabk \theta$) originate from the non-abelian Berry connection. The first term of  $\calh$ corresponds to the spin-orbit-induced splitting of the energy-degenerate bands;
the remaining terms may be interpreted as B-induced deformations of the bands.   It is well-known that the Zeeman term tends to polarize spin in the direction ($\vec{z}$) of $\bB$; remarkably, the non-abelian Berry term   also has a polarizing effect, but with opposite sign. These two effects completely cancel for free electrons ($m{=}m_0$, with $m_0$ being free electron mass.), but not generically in band systems. The orbital moment term $B M_z$, which generally couples the two-band subspace (of interest) to all other bands, is absent in any simplified two-band model.
%Observe that  that $\calh$ is actually time-independent, which originates from a  that is unbroken by the field (aligned with the rotational axis). Consequently,

@@

Our quantization rule equally has utility where no closed-form analytic expression exists for the spectrum. Such is the case when we introduce a Dresselhaus spin-orbit coupling:
\begin{equation}
 H_{RD}(\bk)=\frac{{\hbar^2}k^2}{2m}+\alpha(k_{x}\sigma_{y}-k_{y}\sigma_{x})+\beta(k_{x}\sigma_{x}-k_{y}\sigma_{y}).\la{hamRD}
\end{equation}
In Fig. \ref{fig:RD}(b), a comparison is made between the Landau levels obtained from exact numerical diagonalization of $H_{RD}(\bK)$ (by standard techniques reviewed in App. \ref{sec:numerical}) and those obtained from the quantization rule. Perfect match between quantization rule and exact diagonalization can be observed in the left panel of \fig{fig:RD}(b) while slight deviations are seen for smaller magnetic field. In \fig{fig:RD}(c) we plot the errors in energies for various choices of the spin-orbit parameters $(\alpha,\beta)$, which does not exceed $2\%$ of $\var_c$ in the plot range. Larger errors are positively correlated with 
$l^2\delta S^2/S$ (with $\delta S$ depending on both $\alpha$ and $\beta$), as shown in Fig. \ref{fig:RD}(d); see also \q{eq:Rashba-exact}, which analytically demonstrates how terms involving $l^2\delta S^2/S$ modifies the quantization rule in the case $\beta=0$.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the Hamiltonian $H_{RZ}{=}d_0(\bk)I{+}d_x(\bk)\sigma_x{+}d_y(\bk)\sigma_y$, with $d_j$ obtainable from \q{eq:RZ-Hamiltonian}. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-orbit field. (b) Black dashed lines illustrate a zeroth-order approximation of the Landau levels (assuming Klein tunneling); the leading-order correction is indicated by red lines. The Dirac point lies at $10$meV. Parameters for (b): $m$: 0.02$\hbar^2$/eV$\cdot$\AA$^2$, $\alpha$: 0.05eV$\cdot$\AA, $Z_\parallel$: 0.001eV. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4<\epsilon_0<1.0$(eV).\cite{RZfignote} Parameters for (d): $m$: 0.01$\hbar^2$/eV$\cdot$\AA$^2$, $l=100$\AA. \label{fig:RZ}}
\end{figure}

@@

As alluded to in \s{sec:qtznrules}, the adiabatic limit of two independent orbits does not exist if the two spin-orbit-split orbits exactly intersect at a II-Dirac point. To investigate this failure of adiabaticity, let us study the Landau levels at energies close to the II-Dirac point ($\epsilon_0$). We focus first on the regime where, on average (over $\frako_0$), the energy splitting $|\delta \var_+{-}\delta \var_-|$ (due to both spin-orbit and in-plane Zeeman interactions) is much greater than the interband, out-of-plane Zeeman interaction $B_{\perp}|\calm_{\pm}|$. This implies that the split orbits do not appreciably hybridize -- except in the vicinity of the II-Dirac point where orbits (nearly) touch. In the constant basis, 
\begin{equation}
\calh(t)=\alpha(k_{x}\sigma_{y}-k_{y}\sigma_{x})+Z_\parallel\sigma_{y}-Z_\perp\sigma_{z},\label{calHRZ}
\end{equation}
where $k_x=-k_F \cos(\omega_c t)$ and $k_y=k_F \sin(\omega_c t)$. Near the II-Dirac point, the dynamics of a two-component wavepacket is described by linearizing $\H$ with respect to $t$
\e{
\calh = \alpha k_{E} \,\omega^{\sma{\perp}}_c t\,\sigma_x+(Z_\parallel-\alpha k_{E})\sigma_y-Z_\perp\sigma_z,\la{effhamIIDirac}
}
with $k_E{=}\sqrt{2mE}/\hbar$ and  $t$ a time variable which parametrizes the zeroth-order orbit:
$\bk(t){=}({-}k_E \cos \omega_c^{\sma{\perp}} t, k_E \sin \omega_c^{\sma{\perp}} t ).$ 
The eigenbasis of $\sigma_x$ corresponds to two diabatic levels with characteristic slope $v_d{:}{=}\alpha k_E\omega^{\sma{\perp}}_c$; near $t{=}0$, the diabatic levels anticross  due to the hybridization terms proportional to $\sigma_{y,z}$. Generally, the probability of tunneling across the hybridization-induced barrier is  $\rho^2{=}\exp(-2\pi\barmu)$, with $\barmu{=}{E_g}^2/{2v_d \hbar}$ and $E_g$ the energy gap at the center of the anticrossing;[Landau and lifshitz, Wittig] in this context,
\e{\barmu=\frac{Z_\perp^2+(Z_\parallel-\alpha k_{E})^2}{2\alpha k_{E}\var^{\sma{\perp}}_c}, \la{mu1}}
with $2\alpha k_E$ the spin-orbit splitting at energy $E$ [cf.\ \fig{fig:orbits}(a)]. In the absence of the out-of-plane Zeeman coupling ($Z_\perp{=}0$), $\bar{\mu}$ would vanish at the energy of the II-Dirac point -- leading to unit-probability Klein tunneling, as was first proposed by [Obrien] in a different model realization of the II-Dirac point. However, with a proper account of $Z_\perp$, we find instead that $\bar{\mu}$ has a nonvanishing minimum: $\bar{\mu}_{\sma{\text{min}}}{:}{=} ( m/m_0)(Z_\perp/4\alpha k_{E})$, which is a product of the dimensionless effective mass and the ratio of the out-of-plane Zeeman splitting over the spin-orbit splitting. This minimum is practically negligible for  semiconductor heterostructures with $m{\ll}m_0$, but may be relevant in heavy-fermion systems.

To determine the Landau levels, one needs not just the probability for tunneling but also the phase of the probability amplitude. The full information is encoded in a scattering matrix [AALG]
\e{\mathbb{S}=\matrixtwo{\tau e^{i\barphi}}{-\rho}{\rho}{\tau e^{-i\barphi}},\;\rho=e^{-\pi\barmu},\;\tau=\sqrt{1-\rho^{2}}\la{scattmat}}
which is a connection formula that matches the two-component semiclassical (WKB) wavefunction across the Dirac point. $\mathbb{S}$ in \q{scattmat} has been expressed in the basis of Zeeman-corrected energy bands, i.e., eigenvectors of Eq. (\ref{calHRZ}), so the diagonal elements of $\mathbb{S}$ are amplitudes for intraband (adiabatic) transmission; the phase of this amplitude is $\barphi{:}{=}\barmu{-}\barmu\ln\barmu{+}\text{arg}[\Gamma(i\barmu)]{+}\pi/4$ with $\Gamma$ the gamma function.

Away from the II-Dirac point, a semiclassical wavepacket undergoes adiabatic (band-conserving) dynamics. The time-ordered, coherent process of interband tunneling and adiabatic dynamics is described by the propagator:
\e{{\A}= \matrixtwo{\tau e^{i\bar{\varphi}}}{-\rho}{\rho}{\tau e^{-i\bar{\varphi}}} \diagmatrix{e^{i\tilde{\lambda}_+}}{e^{i\tilde{\lambda}_-}}. \la{propinplanezeeman}}
In the basis of Zeeman-corrected energy bands, the adiabatic phases $\tilde{\lambda}_{\pm}$ consist of two parts: dynamical phase $l^2 \delta S'$ with $\delta S'$ now accounting for both spin-split and Zeeman-split area corrections, and single band Berry phases $\phi^B_\pm$ evaluated on $\frako_0$. At energies far above the Dirac point ($|E{-}\epsilon_0|{\gg}|Z_{\perp}|$) where the Zeeman interaction is negligible, both orbits encircle the Dirac point and hence $\phi_{\pm}^B{\approx}{\pm}\pi$; far below the Dirac point, neither orbit encircles the Dirac point, leading to $\phi_{\pm}^B{\approx}0$. At intermediate energies, $\phi_{+}^B$ (resp.\ $\phi_{-}^B$) is a continuous and increasing (resp.\ decreasing) function of energy, as supported by the Bloch-sphere argument in Fig. \ref{fig:blochsphere}.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{blochsphere.png}
\caption{Energy dependence of the Berry phases for inner (upper panel) and outer (lower panel) orbits. For two band models, the Berry phase equals half the solid angle subtended by the loop swept by spin-splitting field ($-\alpha k_y$,$Z_\parallel+\alpha k_x$,$Z_\perp$) of $\H$ in Eq. (\ref{calHRZ}) on the unit sphere. Notice that Zeeman coupling are included in spin-splitting field here. For lower band (outer orbit), the spin splitting field should be inverted to reflect the change in band index. The solid angles are indicated by blue outlines and the energy is measured from $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}).\label{fig:blochsphere}}
\end{figure}

Our quantization rule [\qq{eq:rule}{eq:H1}], with $\A$ having the form of \q{propinplanezeeman}, may be equivalently expressed as
\e{
\text{cos}\left[\frac{\Omega_{-}+\Omega_{+}}{2}\right]=\tau\,\text{cos}\left[\frac{\Omega_{-}-\Omega_{+}}{2}+\bar{\varphi}\right], \lin
\ins{with}\;\Omega_{\pm}=l^2 (S\pm \delta S'/2)+\phi_{\pm}^B +\gamma_{\pm}, \la{coscos}
}
$\Omega_{\pm}$ is the phase acquired by an electron in adiabatically traversing orbit $\pm$; the quantum Maslov correction $\gamma_{\pm}{=}\pi$ for circular orbits. Since $\phi^B_\pm$ are evaluated both on $\frako_0$, $\phi^B_-+\phi^B_+=0$ in two band models. A quantization rule analogous to \q{coscos} was previously derived by us for a II-Dirac point that intermediates an electron- and hole-like orbit [AALG].
